---
title: "pracice1_staa"
author: "dawnstaa"
date: "1/26/2022"
output: html_document
---


### Reading the data for Antarctica and Greeenland ice loss

```{r setup, include=FALSE}
# Antarctica ice mass loss, Gt - gigatonnes
ant_ice_loss = read.table("data/antarctica_mass_200204_202111_staa.txt", skip=31, sep="", header = FALSE, col.names = c("decimal_date", "mass_Gt", "sigma_Gt")) #last part manually inserts column names after reading header
# Greenland ice mass loss, Gt - gigatonnes
grn_ice_loss = read.table("data/greenland_mass_200204_202111_staa.txt", skip=31, sep="", header = FALSE, col.names = c("decimal_date", "mass_Gt", "sigma_Gt"))
```

### Viewing data and getting general info abt data

```{r}
# View(grn_ice_loss)
head(ant_ice_loss) #returned the first 6 rows of the table/data frame
tail(ant_ice_loss)#returned the last 6 rows of the table/data frame
head(grn_ice_loss)
tail(grn_ice_loss)#returned the last 6 rows of the table/data frame
dim(ant_ice_loss)
summary(ant_ice_loss)
summary(grn_ice_loss)
```

### plotting data using 'y ~ x'

```{r}
# Plot Antarctic ice mass loss
plot(mass_Gt ~ decimal_date, data=ant_ice_loss, ylab="Antarctica Mass Loss (Gt)")
# Plot Greenland ice mass loss
plot(mass_Gt ~ decimal_date, data=grn_ice_loss, ylab="Greenland Mass Loss (Gt)")
# Plot both together as lines
plot(mass_Gt ~ decimal_date, data=ant_ice_loss, ylab="Ice Sheet Mass Loss (Gt)", type='l') 
lines(mass_Gt ~ decimal_date, data=grn_ice_loss, col='cyan') 
```


### Using '+' to link multiple plots and print to the same window

```{r}
plot(mass_Gt ~ decimal_date, data=ant_ice_loss, ylab="Antarctica Mass Loss (Gt)", type='l', ylim=range(grn_ice_loss$mass_Gt)) +
lines(mass_Gt ~ decimal_date, data=grn_ice_loss, type='l', col='cyan') 
```

### Plotting to account for years with missing data
#### insert an `NA` into the data.frame at the start of 2018. Then add in grey shaded polygons behind it to show the uncertainty

```{r}
#Create data.frame with an NA between the GRACE missions. Column names must match so it will merge with ant and grn data
data_break = data.frame(decimal_date=2018.0, mass_Gt=NA, sigma_Gt=NA)
data_break

#Add NA data point to the Antarctica ice trends data frame
ant_ice_loss_with_NA = rbind(ant_ice_loss, data_break) # Merge ant_ice_loss data frame with our NA point
tail(ant_ice_loss_with_NA) # Our NA value in 2018 is now the last row of our data frame
order(ant_ice_loss_with_NA$decimal_date) # Use order() to list the indices of the data after its sorted from smallest to biggest
ant_ice_loss_with_NA = ant_ice_loss_with_NA[order(ant_ice_loss_with_NA$decimal_date),] # Reorder the data frame, sorting by date

#Repeating with Greenland data.frame
grn_ice_loss_with_NA = rbind(grn_ice_loss, data_break) # Merge grn_ice_loss data frame with our NA point
grn_ice_loss_with_NA = grn_ice_loss_with_NA[order(grn_ice_loss_with_NA$decimal_date),]

#Replotting to show the gap
plot(mass_Gt ~ decimal_date, data=ant_ice_loss_with_NA, ylab="Antarctica Mass Loss (Gt)", type='l', ylim=range(grn_ice_loss_with_NA$mass_Gt, na.rm=TRUE)) +
lines(mass_Gt ~ decimal_date, data=grn_ice_loss_with_NA, type='l', col='cyan') 
```
### Plotting C.I around data
#### Sometimes we show errors as "2-sigma", where 2 X sigma represents 95% uncertainty. We can plot 95% confidence intervals around the data by adding and subtracting 2*sigma from the mass loss estimates
```{r}
#Antarctica
head(ant_ice_loss_with_NA)
plot(mass_Gt ~ decimal_date, data=ant_ice_loss_with_NA, ylab="Antarctica Mass Loss (Gt)", xlab="Year", type='l') +
lines((mass_Gt+2*sigma_Gt) ~ decimal_date, data=ant_ice_loss_with_NA, type='l', lty='dashed') +
lines((mass_Gt-2*sigma_Gt) ~ decimal_date, data=ant_ice_loss_with_NA, type='l', lty='dashed') # use 'lines' instead of plot if you want to keep plotting on the same window/plot
```
### Exercise 1.1 Creating the above C.I for Greenland

```{r}
#Greenland Mass loss with 95% confidence interval
head(grn_ice_loss_with_NA)
plot(mass_Gt ~ decimal_date, data=grn_ice_loss_with_NA, ylab="Greenland Mass Loss (Gt)", xlab="Year", type='l') +
lines((mass_Gt+2*sigma_Gt) ~ decimal_date, data=grn_ice_loss_with_NA, type='l', lty='dashed') +
lines((mass_Gt-2*sigma_Gt) ~ decimal_date, data=grn_ice_loss_with_NA, type='l', lty='dashed') 
```

### Combining Antarctic and Greenland plots into one and saving

```{r}
pdf('figures/ice_mass_trends.pdf', width=7, height=5)
plot(mass_Gt ~ decimal_date, data=ant_ice_loss_with_NA, ylab="Ice Sheet Mass Loss (Gt)", xlab="Year", type='l', ylim=range(grn_ice_loss_with_NA$mass_Gt, na.rm=TRUE), lwd=2) +
lines((mass_Gt+2*sigma_Gt) ~ decimal_date, data=ant_ice_loss_with_NA, type='l', lty='dashed') +
lines((mass_Gt-2*sigma_Gt) ~ decimal_date, data=ant_ice_loss_with_NA, type='l', lty='dashed') +
lines(mass_Gt ~ decimal_date, data=grn_ice_loss_with_NA, type='l', col='cyan', lwd=2) +
lines((mass_Gt+2*sigma_Gt) ~ decimal_date, data=grn_ice_loss_with_NA, type='l', lty='dashed', col="cyan") +
lines((mass_Gt-2*sigma_Gt) ~ decimal_date, data=grn_ice_loss_with_NA, type='l', lty='dashed', col="cyan") 
dev.off()
```
### Bar plot showing the difference in ice mass loss for Antarctica and Greenland

```{r}
# Largest observed decrease in ice mass loss in Antarctica:
min(ant_ice_loss$mass_Gt, na.rm =TRUE)
min(grn_ice_loss$mass_Gt, na.rm = TRUE)
# Barplot of largest observed ice loss in Antarctica and Greenland
barplot(height=c(min(ant_ice_loss$mass_Gt), min(grn_ice_loss$mass_Gt)))

# Flip to negative to positive, add x-axis labels, add more tick marks on y-axis, add y-axis title
barplot(height=c(min(ant_ice_loss$mass_Gt)*(-1), min(grn_ice_loss$mass_Gt)*(-1)), names.arg=c("Antarctica","Greenland"), ylim=c(0,5000), ylab="Ice loss in Gt")
```

### Exercise 1.2:

Calculating the average annual ice loss (a.k.a. the annual rate of ice loss) for each ice sheet. Then displaying ice loss rates in a bar graph. Saving bar graph into the `figures/` directory in this repo.

```{r}

```
